Utilitarianism

### 2016-01-31 math_modeling/class/notes/classnotes.tex
so much homework done

@@

We can now consider a discreet case for modeling.
\subsection{Credit Card Balancing}
We can work with the equation
\[
B_{1} = B_{0} - P
\]
as a simple model for a credit card balance. We will eventually find that this model is very
simplistic, and we will need to expand on it later. We have a recurrence relation where
\begin{gather*}
B_{2} = B_{1} - P\\
B_{3} = B_{2} - P\\
\cdots
\end{gather*}
Substituting into the recurrence relation tells us that
\[
B_{n} = B_{0} - nP.
\]
We can reform the model to help use come up with a better representation of the situation.
In the real world, credit cards often come with interest rates. Taking this into account
gives us
\[
B_{\text{new}} = (1+r)B_{\text{old}} - P.
\]
We again have a recurrence relation. We will let $s = (1+r)$:
\begin{align*}
B_{1} &= sB_{0} - P\\
B_{2} &= sB_{1} - P = s^{2}B_{0}-p(1+s)\\
B_{3} &= sB_{2} - P = s^{3}B_{0}-p(1+s+s^{2})\\
B_{4} &= sB_{3} - P = s^{4}B_{0}-p(1+s+s^{2}+s^{3})\\
\cdots
\end{align*}
We can derive a general solution of the form
\begin{align*}
B_{n} &= s^{n}B_{0}-p(1 + s + s^{2} + s^{3} + \cdots + s^{n-1})\\
&= (1+r)^{n}B_{0}-p\left({(1+r)^{n} - 1\over r}\right)
\end{align*}
\subsubsection{Solving With Excel}
We can use Excel to represent the model. We will consider the case where $r=1.5\%$ with an initial
balance of $\$1000$, and $\$10$ per month on payments. In the A1 cell, we want to have
\begin{minted}{text}
=(1.015)^A1 * 1000 - 10*((1.015)^A1-1)/0.015
\end{minted}
\end{document}

>>>>

\usepackage{minted}

@@

We can now consider a discreet case for modeling.
\subsection{Credit Card Balancing}
We can work with the equation
\[
B_{1} = B_{0} - P
\]
as a simple model for a credit card balance. We will eventually find that this model is very
simplistic, and we will need to expand on it later. We have a recurrence relation where
\begin{gather*}
B_{2} = B_{1} - P\\
B_{3} = B_{2} - P\\
\cdots
\end{gather*}
Substituting into the recurrence relation tells us that
\[
B_{n} = B_{0} - nP.
\]
We can reform the model to help use come up with a better representation of the situation.
In the real world, credit cards often come with interest rates. Taking this into account
gives us
\[
B_{\text{new}} = (1+r)B_{\text{old}} - P.
\]
We again have a recurrence relation. We will let $s = (1+r)$:
\begin{align*}
B_{1} &= sB_{0} - P\\
B_{2} &= sB_{1} - P = s^{2}B_{0}-p(1+s)\\
B_{3} &= sB_{2} - P = s^{3}B_{0}-p(1+s+s^{2})\\
B_{4} &= sB_{3} - P = s^{4}B_{0}-p(1+s+s^{2}+s^{3})\\
\cdots
\end{align*}
We can derive a general solution of the form
\begin{align*}
B_{n} &= s^{n}B_{0}-p(1 + s + s^{2} + s^{3} + \cdots + s^{n-1})\\
&= (1+r)^{n}B_{0}-p\left({(1+r)^{n} - 1\over r}\right)
\end{align*}
\subsubsection{Solving With Excel}
We can use Excel to represent the model. We will consider the case where $r=1.5\%$ with an initial
balance of $\$1000$, and $\$10$ per month on payments. In the A1 cell, we want to have
\begin{minted}{text}
=(1.015)^$A1 * 1000 - 10*((1.015)^$A1-1)/0.015
\end{minted}

\section{Modeling Ordinary Differential Equation}
Consider the following system of ODE's
\[
\left[
\begin{array}{c}
Q_{1}^\prime\\
Q_{2}^\prime\\
Q_{3}^\prime
\end{array}
\right]
\]
We will use arbitrary values for the coefficient matrix. We want to solve the system using
eigenvalues:
\[
\left[
\begin{array}{c c c}
5-\lambda & 4 & 2\\
4 & 5-\lambda & 2\\
2 & 2 & 2-\lambda
\end{array}
\right]
= 0
\]
We want to take the determinant of the matrix, but first we can do some row reductions:
\[
\left[
\begin{array}{c c c}
1-\lambda & 0 & 0\\
4 & 9-\lambda & 2\\
2 & 4 & 2-\lambda
\end{array}
\right]
\]
Taking that determinant gives us
\[
\lambda = 1 \qquad\qquad\text{and}\qquad\qquad \lambda = 10.
\]

\subsection{Modeling War}
We will model the Nazi party and the Soviet Union military forces. We want to model the rate
at which Soviet Tanks reduce as Nazi anti-tank guns decrease.
\begin{gather*}
\frac{dx}{dt} = -ay\\
\frac{dy}{dt} = -bx
\end{gather*}
In this system, $x'$ represents the rate at which the Soviet tanks decreased, and $y'$ represents
the rate at which German anti-tank guns decrease. We will consider the battle of Kursk, during
which the number of Soviet tanks decreased by $50\%$ in the first hour. In this model, $a$ is
referred to as the anti-tank kill rate, where $b$ is the tank kill rate. We can use separation
of variables yields
\begin{gather*}
{ay^2\over2}+{ay_{0}^{2}\over2} = {bx^2\over2}+{bx_{0}^{2}\over2}\\
ay^{2}-bx^{2}=C
\end{gather*}
\end{document}

### 2016-01-31 number_theory/hw/hw3.tex
so much homework done

\documentclass{hw}
\begin{document}
\makeheader{3}

\begin{enumerate}
\item Compute $a\cdot b$, $||a||$ and $||b||$ where $\vv{a} = 4i -3j + k$ and $\vv{b}=i+j+k$.
\begin{quote}
\begin{align*}
a\cdot b &= 4 - 3 + 1\\
&= 2\\
||a|| &= \sqrt{16+9+1}\\
&= \sqrt{26}\\
||b|| &= \sqrt{1+1+1}\\
&= \sqrt{3}
\end{align*}
\end{quote}

\item Find the angle between $\vv{a} = \sqrt{3}\ i + j$ and $\vv{a} = -\sqrt{3}\ i + j$.
\begin{align*}
\theta &= \arccos{\left({a\cdot b \over ||a||\ ||b||}\right)}\\
&= \arccos{\left({-2\over 4}\right)}\\
&= \arccos{\left(-{1\over 2}\right)}\\
&= {2\pi \over 3}
\end{align*}

\item Find the angle between $\vv{a} = i + j$ and $\vv{b} = i + j + k$.
\begin{align*}
\theta &= \arccos{\left({a\cdot b \over ||a||\ ||b||}\right)}\\
&= \arccos{\left({2 \over \sqrt{6}}\right)}
\end{align*}

\item Find the angle between $\vv{a} = (1,-2,3)$ and $\vv{b} = (3,-6,-5)$.
\begin{align*}
\theta &= \arccos{\left({a\cdot b \over ||a||\ ||b||}\right)}\\
&= \arccos{(0)}\\
&= {\pi \over 2}
\end{align*}

\item Calculate $proj_{a}b$ where $\vv{a} = i + j$ and $\vv{b} = 2i + 3j - k$.
\begin{align*}
proj_{a}b &= {a\cdot b \over ||a||^{2}}\vv{a}\\
&= {4\over 2}(i + j)\\
&= 2i + 2j
\end{align*}

\item Calculate $proj_{a}b$ where $\vv{a} = i + j + 2k$ and $\vv{b} = 2i - 4j + k$.
\begin{align*}
proj_{a}b &= {a\cdot b \over ||a||^{2}}\vv{a}\\
&= {0\over 4}(i+j+2k)\\
&= 0
\end{align*}

\item Give a vector of length 3 that points in the same direction as the vector $i + j - k$.
\begin{quote}
The length is just a scalar value multiplied by the vector. Therefore we have the unit vector $i + j - k$,
with length 3 will be
\[3i + 3j - 3k.\]
\end{quote}

\item Is there ever a case where $proj_{a}b = proj_{b}a$? If so, under what conditions?
\begin{quote}
If we use the explicit formulas equal to each other, we have
\begin{align*}
\left({a\cdot b \over a\cdot a}\right)\vv{a} &= \left({a\cdot b \over b\cdot b}\right)\vv{b}\\
{||a|| \over ||b||}\vv{a} &= \vv{b}
\end{align*}
Therefore, the projections are equal only when $\vv{a} = \vv{b}$.
\end{quote}

\item Prove part 1 of Proposition 3.4.
\begin{quote}
We need to show that $||k\vv{a}|| = |k|\ ||\vv{a}||$. Then
\begin{align*}
||k\vv{a}|| &= ||(ka_{1},ka_{2},\dots,ka_{n})||\\
&= \sqrt{k^{2}a_{1}^{2} + k^{2}a_{2}^{2} + \cdots + k^{2}a_{n}^{2}}\\
&= \sqrt{k^{2}(a_{1}^{2} + a_{2}^{2} + \cdots + a_{n}^{2})}\\
&= \sqrt{k^{2}}\ \sqrt{a_{1}^{2} + a_{2}^{2} + \cdots + a_{n}^{2}}\\
&= |k|\ ||\vv{a}||.
\end{align*}
$\triangle$
\end{quote}

\item In physics, when a constant force acts on an object as the object is displaced, the work done by
the force is the product of the length of the displacement and the component of the force in the
direction of the displacement. Figure 1.48 depicts an object acted upon by a constant force $F$, which
displaces it from the point $P$ to the point $Q$. Let $\theta$ denote the angle between $F$ and the
direction of
displacement.
\begin{enumerate}
\item Show that the work done by $F$ is determined by the formula $F\cdot\vv{PQ}$.
\begin{quote}
We can determine the $x$ component of $F$ as $||F||\cos{\theta}$. Then the work can be computed as
$||F||\ ||\vv{PQ}||\cos{\theta}$, which is $F\cdot\vv{PQ}$ by definition.
\end{quote}
\item Find the work done by the (constant) force $F = i + 5j + 2k$ in moving a particle from the point
$(1, -1, 1)$ to the point $(2, 0, -1)$.
\begin{quote}
We need to find the dot product $(1,5,2)\cdot(1,1,-2)$, which is equal to $2$. Therefore, the work done
is 2 units.
\end{quote}
\end{enumerate}

\item Using vectors, prove that the diagonals of a parallelogram are perpendicular if and only if the
parallelogram is a rhombus. (Note: A rhombus is a parallelogram whose four sides all have the same
length).
\begin{quote}
We need to show that the vectors $\vv{ab}$ and $\vv{ba}$ have a dot product of 0. We know that the
lengths of $\vv{a}$ and $\vv{b}$ are equal, and that $\vv{ab} = (a_{1} - b_{1}, a_{2} - b_{2})$ and
$\vv{ba} = (b_{1} + a_{1}, b_{2} + a_{2})$. Therefore
\begin{align*}
\vv{ab}\cdot\vv{ba} &= (a_{1} - b_{1})(b_{1} + a_{1}) + (a_{2} - b_{2})(b_{2} + a_{2})\\
&= a_{1}^{2}+a_{2}^{2}-b_{1}^{2}-b_{2}^{2}\\
&= b_{1}^{2}+b_{2}^{2}-b_{1}^{2}-b_{2}^{2},\text{ Since the lengths are equal,}\\
&= 0
\end{align*}
\end{quote}
\end{enumerate}
\end{document}

>>>>

\documentclass{hw}

\begin{document}
\makeheader{3}

\begin{enumerate}
\item 
\end{enumerate}
\end{document}

### 2016-01-31 vector_calc/homework/hw6.tex
so much homework done

\begin{document}
\makeheader{6}

@@

\item Find a set of cylindrical coordinates of the point whose Cartesian coordinate is given.
\begin{quote}
\begin{align*}
(-1,\sqrt{3},13) &= (\sqrt{(-1)^2 + (\sqrt{3})^2},\arctan{-{1\over\sqrt{3}}},13)\\
&= (2,-{\pi\over6},13)
\end{align*}
\end{quote}
\end{enumerate}
\end{document}

>>>>

\begin{document}
\makeheader{1.7}

@@

\item Find a set of cylindrical coordinates of the point whose Cartesian coordinate is given.
\begin{quote}
\begin{align*}
(-1,\sqrt{3},13) &= (\sqrt{(-1)^2 + (\sqrt{3})^2},\arctan{-{1\over\sqrt{3}}},13)\\
&= (2,-{\pi\over6},13)
\end{align*}
\end{quote}

\item Convert the provided equation into the other two coordinate systems.
\[
\rho\sin{\phi}\sin{\theta} = 2
\]
\begin{quote}
In cylindrical coordinates: \[r\sin{\theta} = 2\]
In cartesian coordinates: \[y=2\]
\begin{gather*}
\end{gather*}
\end{quote}

\item Convert the provided equation into the other two coordinate systems.
\[
z^{2} = 2x^{2} + 2y^{2}
\]
\begin{quote}
In cylindrical coordinates:
\begin{gather*}
z^{2} = r^{2}\cos^{2}{\theta} + r^{2}\sin^{2}{\theta}\\
z^{2} = r^{2}
\end{gather*}
In spherical coordinates:
\begin{gather*}
\rho^{2}\cos^{2}{\phi} = \rho^{2}\sin^{2}{\phi}\cos^{2}{\theta} +
\rho^{2}\sin^{2}{\phi}\sin^{2}{\theta}\\
\cos^{2}{\phi} = \sin^{2}{\phi}(\cos^{2}{\theta} + \sin^{2}{\theta})\\
\cos^{2}{\phi} = \sin^{2}{\phi}
\end{gather*}
\end{quote}

\item Sketch the solid whose cylindrical coordinates $(r,\theta,z)$ satisfy the given
inequalities.
\[
0\leq r\leq3,\ 0\leq\theta\leq{\pi\over2},\ -1\leq z\leq2
\]
\vspace{3cm}

\item Sketch the solid whose cylindrical coordinates $(r,\theta,z)$ satisfy the given
inequalities.
\[
r\leq z\leq 5,\ 0\leq\theta\leq\pi
\]
\vspace{3cm}

\item Sketch the solid whose spherical coordinates satisfy the given inequalities.
\[
0\leq\rho\leq1,\ 0\leq\phi\leq{\pi\over2}
\]
\vspace{3cm}

\item Sketch the solid whose spherical coordinates satisfy the given inequalities.
\[
0\leq\phi\leq{\pi\over4},\ 0\leq\rho\leq2
\]
\vspace{3cm}

\item Consider the solid shown below:
\begin{center}
\includegraphics[scale=0.4]{hw6_pic}
\end{center}
\begin{enumerate}
\item Describe the solid, using spherical coordinates.
\[
0\leq\rho\leq3,\ 0\leq\phi\leq\arctan{1\over\sqrt{8}},\ 0\leq\theta\leq2\pi
\]
\item Describe the solid, using cylindrical coordinates.
\[
\sqrt{8}\ r\leq z\leq\sqrt{9-r^{2}},\ 0\leq r\leq 1,\ 0\leq\theta\leq2\pi
\]
\end{enumerate}
\end{enumerate}
\end{document}

### 2016-01-31 writing_in_ethics/paper_one.tex
so much homework done

\begin{document}
\drafth{}
\inlinetitle
\begin{linenumbers}
The theory of Utilitarianism can be used to quantitatively analyze various situations that arise in day to day life. It can be used to analyze the scenario presented in \textit{Ethics for the Information Age} in which a promising high school student, Alexis, steals login information in order to use a computer to which she does not have access. Alexis' situation can be analyzed under both Rule and Act Utilitarianism, which both require a numerical value to be computed based on the actions of an individual. In Alexis' case, there is no clear unit that can be assigned to the value of her actions; rather, the overall general outcome can be used to gauge the utility of Alexis' actions.

Jeremy Bentham, a famous English philosopher, proposed several attributes that should be taken into account when calculating the utility of an action \cite[75]{ethics}. The first attribute, \textit{Intensity}, is concerned with the magnitude of the experience. The magnitude of Alexis' experience can be measured in the amount of happiness she experienced. She was able to achieve her goal and even positively affected others around her; presumably she was able to save her family and herself a large amount of money while also bettering her own life in the process through a college education. Both of these aspects should be taken into account as contributing positively to the magnitude of the situation.

The next attribute, \textit{Duration}, applies greatly to Alexis' scenario. The duration of the experience extends far beyond the time that Alexis spent in the library on the computer; in fact, the education she received will continue to affect her for the rest of her life. However, Alexis stole the information from another student at the university. The duration of that action, coupled with the amount of time she spent using the stolen information, is still greatly overshadowed by the overall happiness that Alexis experienced. It is reasonable to say that the amount of happiness gained under the \textit{Duration} criterion is far greater than the amount of happiness lost.

Bentham mentions the idea of certainty in his criteria for testing actions under Act Utilitarianism. There was always a chance that Alexis' end goal could have never been met, in which case the overall unhappiness produced would have far outweighed the happiness. However, Alexis did ultimately achieve her goal to be accepted into college; on top of just being accepted into college, she was also provided with a full ride, which positively affected her parents as well. As a result, the attribute of \textit{Certainty} greatly adds to the overall happiness gained from Alexis' decision.

The final attribute that Bentham proposes is concerned with the number of people affected by the action. Alexis' actions do in fact affect several other beings, and not all the interactions are not necessarily positive. As mentioned before, Alexis' family benefitted greatly from the situation. They were already poor; this fact may have greatly affected Alexis' ability to attend college. However, due to her actions, Alexis was able to obtain a scholarship that provided her with the funds to attend college for free. In this way, Alexis affected, in a positive way, a large number of people that were important to her. However, she did also affect the student from whom she stole the login information. Her actions had a slight negative affect on that student, but not enough of a negative affect to outweigh the positive affect she had on her family.

Bentham's methods applied to Alexis' actions under the ethical theory of Act Utilitarianism. Under Act Utilitarianism, Alexis has made an ethically good decision; the amount of overall happiness produced by her actions far outweighs the unhappiness produced. However, many of the calculations are not exact since the certainty of the consequences of Alexis' actions were often assumed \cite[77]{ethics}. The theory of Rule Utilitarianism can also provide a viewpoint to determine whether or not Alexis' actions were ethically right or wrong.

>>>>

\begin{document}
\finalh{}
\inlinetitle
\begin{linenumbers}

The theory of Utilitarianism can be used to quantitatively analyze various situations that arise in day to day life. One can analyze the scenario presented in \textit{Ethics for the Information Age} in which a promising high school student, Alexis, steals login information in order to use a computer to which she does not have access. Alexis' situation can be analyzed under both Rule and Act Utilitarianism, which both require a numerical value to be computed based on the actions of an individual. In Alexis' case, there is no clear unit of measurement that can be assigned to the value of her actions; rather, the overall general outcome can be used to gauge the utility of Alexis' actions.

Jeremy Bentham, a famous English philosopher, proposed several attributes that should be taken into account when calculating the utility of an action \cite[75]{ethics}. The first attribute, \textit{Intensity}, is concerned with the magnitude of the experience. The magnitude of Alexis' experience can be measured in the amount of happiness she experienced. She was able to achieve her goal and even positively affect others around her. Presumably she was able to save her family and herself a large amount of money while also bettering her own life through a college education. Both of these aspects should be taken into account as contributing positively to the overall utility of the situation.

The next attribute, \textit{Duration}, applies greatly to Alexis' scenario. The duration of the experience extends far beyond the time that Alexis spent in the library on the computer. In fact, the education she received will continue to affect her for the rest of her life. However, Alexis stole the information from another student at the university. The duration of that action, coupled with the amount of time she spent using the stolen information, is still greatly overshadowed by the overall happiness that Alexis experienced. It is reasonable to say that the amount of happiness gained under the \textit{Duration} criterion is far greater than the amount of happiness lost.

Bentham mentions the idea of certainty in his criteria for testing actions under Act Utilitarianism. There was always a chance that Alexis' end goal could have never been met, in which case the overall unhappiness produced would have far outweighed the happiness. However, Alexis did ultimately achieve her goal to be accepted into college. In addition to just being accepted into college, she was also provided with a full ride, which positively affected her parents as well. As a result, the attribute of \textit{Certainty} greatly adds to the overall happiness gained from Alexis' decision.

The final attribute that Bentham proposes is concerned with the number of people affected by the action. Alexis' actions do in fact affect several other beings, and not all the interactions are necessarily positive. As mentioned before, Alexis' family benefitted greatly from the situation. They were already poor; this fact may have greatly affected Alexis' ability to attend college. However, due to her actions, Alexis was able to obtain a scholarship that provided her with the funds to attend college for free. In this way, Alexis affected, in a positive way, a large number of people that were important to her. However, she did also affect the student from whom she stole the login information. Her actions had a slight negative affect on that student, but not enough of a negative affect to outweigh the positive affect she had on her family.

Bentham's methods also apply to Alexis' actions under the ethical theory of Act Utilitarianism. Under Act Utilitarianism, Alexis has made an ethically good decision; the amount of overall happiness produced by her actions far outweighs the unhappiness produced. However, many of the calculations are not exact since the certainty of the consequences of Alexis' actions were often assumed \cite[77]{ethics}. The theory of Rule Utilitarianism can also provide a viewpoint to determine whether or not Alexis' actions were ethically right or wrong.
